### 2024-11-23 paper/intro.tex
Minor changes
WIP: write al
Minor change

% Rigorous standardization -> spec is rigorous: formal notation & prose notation
To address this challenge, Wasm has been rigorously standardized by the
W3C~\cite{wasm-w3c}.
The Wasm specification is particularly rigorous, describing its semantics in
two complementary forms: \textit{formal notation} and \textit{prose notation}.
\cref{fig:testop} illustrates the specification document for $testop$
instruction, where the upper part contains the prose notation, and the lower
part contains the formal notation.
The formal notation employs mathematical rules to succinctly define the
semantics, supporting proofs such as type soundness.
Conversely, the prose notation uses pseudocode-like algorithms to explain the
semantics through step-by-step instructions.
Since most Wasm users and engine developers are less familiar with mathematical
formalism, they primarily rely on the prose notation.

@@

% Challenging specification process
The demanding standardization process places a significant burden on
specification authors, particularly in maintaining consistency and correctness.
Crafting the specification document is labor-intensive, which becomes
increasingly difficult to scale as Wasm evolves.
This challenge is compounded by the dual requirement to maintain both a formal
notation, written in LaTeX, and a prose notation, authored in reStructuredText.
The separate and user-unfriendly nature of these formats hinders collaborative
review and increases the risk of discrepancies and errors between the two
notations, further complicating the standardization process.

@@

% Challenge
One challenge of our approach is making AL executable.
Since the prose notation in the official specification is written in informal
pseudocode, the interpretation of the prose notation is not straightforward.
The naive implementation of AL interpreter produced incorrect results when
running Wasm official test suites, particularly those related to control flow.
Additionally, the absence of a formalized definition or formal semantics for AL
made it difficult to identify and address the fundamental issues in its design.
Since running tests is the primary method for assessing the correctness of the
specification, this limitation hindered the development of both the generated
interpreter and the specification itself.
This underscored the need for a more robust approach to formally define AL and
ensure its alignment with Wasm's complex semantics.

@@

% TODO: cref
% Contribution
The contributions of this paper are as follows:
\begin{itemize}
  \item We propose a formal model for Wasm control flow in AL (\cref{sec:model})
  \item We formalize syntax and semantics of AL (\cref{sec:formal})
  \item We implement an AL interpreter based on the formalization (\cref{ch:interpreter})
\end{itemize}

>>>>

% Rigorous standardization -> formal notation & prose notation
To address this challenge, Wasm has been rigorously standardized by the
W3C~\cite{wasm-w3c}.
The Wasm specification is particularly rigorous, describing its semantics in
two complementary forms: \textit{formal notation} and \textit{prose notation}.
\cref{fig:testop} illustrates the specification document for $testop$
instruction, where the upper part contains the prose notation, and the lower
part contains the formal notation.
The formal notation employs mathematical rules to succinctly define the
semantics, supporting proofs such as type soundness.
Conversely, the prose notation uses pseudocode-like algorithms to explain the
semantics through step-by-step instructions.
Since most Wasm users and engine developers are less familiar with mathematical
formalism, they primarily rely on the prose notation.

@@

% Challenging specification process
The demanding standardization process places a significant burden on
specification authors, particularly in maintaining consistency and correctness.
Crafting the specification document is labor-intensive, which becomes
increasingly difficult to scale as Wasm evolves.
This challenge is compounded by the dual requirement to maintain both a formal
specification, written in LaTeX, and a prose specification, authored in
reStructuredText.
The separate and user-unfriendly nature of these formats hinders collaborative
review and increases the risk of discrepancies and errors between the two
notations, further complicating the standardization process.

@@

% Challenge
One challenge of our approach is making AL executable.
Since the prose specification is written in informal pseudocode, its
interpretation is not straightforward.
The naive implementation of AL interpreter produced incorrect results when
running Wasm official test suites, particularly those related to control flow.
Additionally, the absence of a formalized definition or formal semantics for AL
made it difficult to identify and address the fundamental issues in its design.
Since running tests is the primary method for assessing the correctness of the
specification, this limitation hindered the development of both the generated
interpreter and the specification itself.
This underscored the need for a more robust approach to formally define AL and
ensure its alignment with Wasm's complex semantics.

@@

% Contribution
The contributions of this paper are as follows:
\begin{itemize}
  \item We propose a formal model for Wasm control flow in AL
    (\cref{sec:model})
  \item We formalize syntax and semantics of AL (\cref{sec:formal})
  \item We implement an AL interpreter based on the formalization
    (\cref{ch:interpreter})
\end{itemize}
